---
title: "ZIP"
author: "Shan Chen"
date: "5/4/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = F) 
suppressMessages(library(tidyverse))
suppressMessages(library(factoextra))
suppressMessages(library(ggrepel))
```
#Eigenspace
```{r}
dataDir <- "/Users/shawnchen/Desktop"
test.df <- read.csv(file.path(dataDir,"zipTestSmall.csv"),header=T) 
train.df <- read.csv(file.path(dataDir,"zipTrainSmall.csv"),header=T)
```

```{r}
zipTest <- list()
zipstrain <- list()
for(i in 1:10){
  zipTest [[i]] <- test.df %>%
    filter(V1==i-1) %>%
    select(-V1) %>%
    t()%>%
    scale(scale = F)
  zipstrain [[i]] <- train.df %>%
    filter(V1==i-1) %>%
    select(-V1) %>%
    t()%>%
    scale(scale = F)
}

```

Note that n= number of observations, an observation  is a pixel
value.  Also, p = number of features, in this case, its a the pixal
value across all the cats!

## Helper functions

In order to work with these images, we need  to be able to view them. To do so, let's create a couple helper functions  that will facilitate going back and forth between flat and square images. 


The first function converts an image into 64x64 matrix.
```{r}
imageConv <- function(flatImage,size=64){
    matrix(flatImage,nrow=size,byrow=T)    
}

```


The second function  reverses the process. 
```{r}
flatConv <- function(rectImage,size=64){
    matrix(t(rectImage),nrow=size^2,ncol=1)[,1]
}

```

This makes it easy to plot our images with a minimum of fuss and muss.
```{r}
flatImage <- function(dat) {
  img <- imageConv(dat)
  image(img,col=grey.colors(256)) 
}
```

# Singular Value Decomposition

Now we can begin the computation in the image space. Remember the plan.

 * Perform the SVD decomposion of the full data set.
 * Identify the eigenvectors that span the column space (subset of image space).  
 * Use the eigenvectors to define the projection mapping from the full image space onto the column space.
 * The closer an image is to a particular eigenspace, the more it is like the common images.

Now the Singular Value Decomposition of the cat images
```{r}
zipTrain.svd <- list()
U <- list()
for(i in 1:10){
zipTrain.svd[[i]] <- svd(zipstrain[[i]])
U[[i]] <- zipTrain.svd[[i]]$u
}
```

## Projections onto space: Eigenface reconstuctions

## 0 Space
Pull off a dog!
```{r}
a0 <- zipstrain[[1]]
a00 <- a0[,1]
```

#10 Eigenspaces

Define all 
```{r}
hat <- list()
for(i in 1:10){
hat[[i]] <- U[[i]] %*% t(U[[i]])
}
```

```{r}
suanyisuan <- function(img){
  proj <- numeric()

  dist <- numeric()
  for(i in 1:10){
    proj[i] <- hat[[i]] %*% img
    dist[i] <- mean((img-proj[i])^2)
  }
  which.min(dist)-1
}

```
    proj[i] <- hat[[i]] %*% zipTest[i1]
Which space is the original image closest to?
#0
```{r}
j=0
for (i in 1:180) {
  jieguo <- suanyisuan(zipTest[[1]][,i])
  if(jieguo!=0){
    j = j +1
  }
}
c(j/180)
```
catDist <- mean((anImage-catProj)^2)
dogDist <- mean((anImage-dogProj)^2)
#1
```{r}
j=0
for (i in 1:129) {
  jieguo <- suanyisuan(zipTest[[2]][,i])
  if(jieguo!=1){
    j = j +1
  }
}
c(j/129)
```
#2
```{r}
j=0
for (i in 1:85) {
  jieguo <- suanyisuan(zipTest[[3]][,i])
  if(jieguo!=1){
    j = j +1
  }
}
c(j/85)
```
#3
```{r}
j=0
for (i in 1:84) {
  jieguo <- suanyisuan(zipTest[[4]][,i])
  if(jieguo!=1){
    j = j +1
  }
}
c(j/84)
```
#4
```{r}
j=0
for (i in 1:128) {
  jieguo <- suanyisuan(zipTest[[5]][,i])
  if(jieguo!=1){
    j = j +1
  }
}
c(j/128)
```
#5
```{r}
j=0
for (i in 1:71) {
  jieguo <- suanyisuan(zipTest[[6]][,i])
  if(jieguo!=1){
    j = j +1
  }
}
c(j/71)
```
#6
```{r}
j=0
for (i in 1:76) {
  jieguo <- suanyisuan(zipTest[[7]][,i])
  if(jieguo!=1){
    j = j +1
  }
}
c(j/76)
```
#7
```{r}
j=0
for (i in 1:74) {
  jieguo <- suanyisuan(zipTest[[8]][,i])
  if(jieguo!=1){
    j = j +1
  }
}
c(j/74)
```
#8
```{r}
j=0
for (i in 1:90) {
  jieguo <- suanyisuan(zipTest[[9]][,i])
  if(jieguo!=1){
    j = j +1
  }
}
c(j/90)
```
#9
```{r}
j=0
for (i in 1:83) {
  jieguo <- suanyisuan(zipTest[[10]][,i])
  if(jieguo!=1){
    j = j +1
  }
}
c(j/83)
```