---
title: "introCV4"
author: "Shan Chen"
date: "04/01/2020"
output: html_document
---

```{r}
library(tidyverse)
library(ISLR) ## for lda
library(MASS)
library(class)#for knn
```

## Assignment 4:  Classification 2
Consider the data set in "ClassificationData2D.csv", use 10-fold cross-validation to determine the best possible KNN predictive model.
```{r}
data2D.df <- read_csv( "ClassificationData2D.csv")
data2D.df %>%
  ggplot()+
  geom_point(aes(x1,x2,color=factor(class)))+
  scale_color_manual(values=c("red","blue"))+
  labs(title="2D Classification Data")


```
KNN：
```{r}
N <- nrow(data2D.df)
p <- ncol(data2D.df)
names(data2D.df)
kVal <- 10
numFolds <- 10
folds <- sample(1:numFolds,N,rep=T)
errs <- numeric(numFolds)
for(fold in 1:numFolds){
  train.x <- data.matrix(data2D.df[folds != fold, 1:2])
  test.x <- data.matrix(data2D.df[folds == fold, 1:2])
  train.y <- data.matrix(data2D.df[folds != fold, 3])
  test.y <- data.matrix(data2D.df[folds == fold, 3])
  mod.knn <- knn(train.x,test.x,train.y,k=kVal)
  errs[fold] <- mean((test.y != mod.knn))
}
mean(errs)
```
knn function
```{r}
errCV.knn <- function(kVal){
  folds <- sample(1:numFolds,N,rep=T)
  errs <- numeric(numFolds)
  for(fold in 1:numFolds){
    train.x <- data.matrix(data2D.df[folds != fold, 1:2])
    test.x <- data.matrix(data2D.df[folds == fold, 1:2])
    train.y <- data.matrix(data2D.df[folds != fold, 3])
    test.y <- data.matrix(data2D.df[folds == fold, 3])
    mod.knn <- knn(train.x,test.x,train.y,k=kVal)
    errs[fold] <- mean((test.y != mod.knn))
  }
  mean(errs)
}
```
run kval over error and plot the result
```{r}
maxK <- 30
kVals <- 2*(1:maxK)+1
errs <- map_dbl(kVals,errCV.knn)
data.frame(k=kVals,err=errs) %>%
  ggplot()+
  geom_point(aes(k,err))+
  geom_line(aes(k,err))
#show optimal choice:
kBest <- 31
(err.knn <- errCV.knn(kBest))
```
How does this  result  compare with Logistic regression, LDA, and QDA.  In each case, just use the full predictor set. Use 10-fold cross-validation to estimate the error rate.
For LDA with 10 folds:
```{r}
folds <- sample(1:numFolds,N,rep=T)
errs <- numeric(numFolds)
for(fold in 1:numFolds){
  train.df <- data2D.df[folds != fold,]
  test.df <- data2D.df[folds != fold, ]
  mod.lda <- lda(class ~ x1+x2,data=train.df)
  pred <- predict(mod.lda,newdata=test.df)
  pred$class
  errs[fold] <- with(test.df,mean((class!=pred$class)))
}
(err.lda <- mean(errs))
```

For QDA with 10 folds:
```{r}
folds <- sample(1:numFolds,N,rep=T)
errs <- numeric(numFolds)
for(fold in 1:numFolds){
  train.df <- data2D.df[folds != fold,]
  test.df <- data2D.df[folds != fold, ]
  mod.qda <- qda(class ~ x1+x2,data=train.df)
  pred <- predict(mod.qda,newdata=test.df)
  pred$class
  errs[fold] <- with(test.df,mean((class!=pred$class)))
}
(err.qda <- mean(errs))
```
error compare:
```{r}
c(err.knn,err.lda,err.qda)
min(err.knn,err.lda,err.qda)
```
KNN is our winner.