---
title: 'IntroCV: Assignment 3- Auto Data'
author: "Shan Chen"
date: "3/31/2020"
output: html_document
---

```{r}
library(tidyverse)
library(ISLR) ## for lda
library(glmnet)
library(class)#for knn
```

## Assignment 3: Classification 1
For the predictive question above using KNN, what is the best value of k? Is it k = 1? Use 10-fold cross-validation. To estimate the best possible
Caution: For some reason, R is knn  is unhappy with k=2.
```{r}
names(Auto)
```

```{r}
c(Auto$year)
with(Auto,table(year))
```
Question: Can we predict which cars were made before 1978?

check data with car config of mpg v acceleration:
```{r}
Auto <- Auto %>%
  mutate(pre78 = year < 78)
Auto%>%
  ggplot()+
  geom_point(aes(mpg,acceleration,color=factor(pre78)))
```
ten folds

```{r}
kVal <- 40
fold <- 1
n <- nrow(Auto)
errCV <- function(kVal,numFolds=10){
  errs <- numeric(numFolds)
  folds <- sample(1:numFolds,n,rep=T)
  for(fold in 1:numFolds){
    train.df <- Auto[folds != fold,]
    test.df <- Auto[folds == fold,]
    train.x <- data.matrix(train.df[c("mpg","acceleration")])
    resp.x<- data.matrix(train.df[c("pre78")])
    test.x <- data.matrix(test.df[c("mpg","acceleration")])
    mod.knn <- knn(train.x,test.x,resp.x,k=kVal)
    errs[fold] <- with(test.df,mean(pre78 != (mod.knn==1)))
  }
  mean(errs)
}
```

As we can see from graph below, it first hits the low around 30 and the lowest hits about 110.
```{r}
## Skip 2 and plot error changes over kval
kMax <- 100
kVals <- 2*(1:kMax)+1
errs <- map_dbl(kVals,errCV)
data.frame(k=kVals,err=errs) %>%
  ggplot()+
  geom_point(aes(k,err))+
  geom_line(aes(k,err))
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


=```{r}
```

