---
title: "LDA and Dimension Reduction"
author: "Shan Chen"
date: "3/8/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(MASS) ## For lda
library(tidyverse)
library(mlbench)

```

```{r}
data("Vehicle")
```
There are four types of cars. See all 4 Class distributions.
```{r}
with(Vehicle,table(Class))
```
```{r}
Vehicle2.df <- Vehicle %>%
filter(Class %in% c("opel","saab")) %>%
mutate(Class01=ifelse(Class=="opel", 1, 0)) %>%
droplevels() ##ignore the other Class factors
```
divide the data to test and train
```{r}
N <- nrow(Vehicle2.df)
train <- sample(1:N,N/2,rep=F)
train.df <- Vehicle2.df[train,]
test.df <- Vehicle2.df[-train,]
mod.log <- glm(Class01~.-Class, data = train.df, family=binomial)
test.df$prob.log <- predict(mod.log, newdata=test.df, type="response")
test.df <- test.df %>%
mutate(Class.log=as.factor(ifelse(prob.log>.5, "opel", "saab")))
with(test.df, table(Class, Class.log))
```

```{r}
(err.log <- with(test.df, mean(Class != Class.log)))
```

```{r}
mod.lda <- lda(Class~.-Class01, data=train.df)
test.df$Class.lda <- predict(mod.lda, newdata=test.df)$class
with(test.df, table(Class, Class.lda))
```

```{r}
(err.lda <- with(test.df, mean(Class != Class.lda)))
```

```{r}
with(test.df, table(Class.log, Class.lda))
```

```{r}
scale.lda <- mod.lda$scaling
(scale.lda <- scale.lda/sqrt(sum(scale.lda^2)))
```

```{r}
data.mat <- data.matrix(
subset(test.df,
select=-c(Class,Class01,prob.log,Class.log,Class.lda)))
proj.vals <- data.mat %*% scale.lda
test.df$proj <- as.numeric(proj.vals)
test.df %>%
ggplot()+
geom_density(aes(proj,fill=factor(Class)))+
labs(title="Separation of Classes(opel and saab) in LDA Space (dimension=1)",
fill="Class",x="")+
facet_grid(Class~.)
```

Other Classes
Repeat the above analysis with “bus” and “van” instead of “opel” and “saab”.
Repeat the above analysis with “opel” and “bus” instead of “opel” and “saab”.

Bus and Van

```{r}
Vehicle2.df <- Vehicle %>%
filter(Class %in% c("bus","van")) %>%
mutate(Class01=ifelse(Class=="bus", 1, 0)) %>%
droplevels() ##ignore the other Class factors
```
divide the data to test and train
```{r}
N <- nrow(Vehicle2.df)
train <- sample(1:N,N/2,rep=F)
train.df <- Vehicle2.df[train,]
test.df <- Vehicle2.df[-train,]
mod.log <- glm(Class01~.-Class, data = train.df, family=binomial)
test.df$prob.log <- predict(mod.log, newdata=test.df, type="response")
test.df <- test.df %>%
mutate(Class.log=as.factor(ifelse(prob.log>.5, "bus", "van")))
with(test.df, table(Class, Class.log))
```

```{r}
(err.log <- with(test.df, mean(Class != Class.log)))
```

```{r}
mod.lda <- lda(Class~.-Class01, data=train.df)
test.df$Class.lda <- predict(mod.lda, newdata=test.df)$class
with(test.df, table(Class, Class.lda))
```

```{r}
(err.lda <- with(test.df, mean(Class != Class.lda)))
```

```{r}
with(test.df, table(Class.log, Class.lda))
```

```{r}
scale.lda <- mod.lda$scaling
(scale.lda <- scale.lda/sqrt(sum(scale.lda^2)))
```

```{r}
data.mat <- data.matrix(
subset(test.df,
select=-c(Class,Class01,prob.log,Class.log,Class.lda)))
proj.vals <- data.mat %*% scale.lda
test.df$proj <- as.numeric(proj.vals)
test.df %>%
ggplot()+
geom_density(aes(proj,fill=factor(Class)))+
labs(title="Separation of Classes(bus and van) in LDA Space (dimension=1)",
fill="Class",x="")+
facet_grid(Class~.)
```

Opel and Bus

```{r}
Vehicle2.df <- Vehicle %>%
filter(Class %in% c("opel","bus")) %>%
mutate(Class01=ifelse(Class=="opel", 1, 0)) %>%
droplevels() ##ignore the other Class factors
```
divide the data to test and train
```{r}
N <- nrow(Vehicle2.df)
train <- sample(1:N,N/2,rep=F)
train.df <- Vehicle2.df[train,]
test.df <- Vehicle2.df[-train,]
mod.log <- glm(Class01~.-Class, data = train.df, family=binomial)
test.df$prob.log <- predict(mod.log, newdata=test.df, type="response")
test.df <- test.df %>%
mutate(Class.log=as.factor(ifelse(prob.log>.5, "opel", "bus")))
with(test.df, table(Class, Class.log))
```

```{r}
(err.log <- with(test.df, mean(Class != Class.log)))
```

```{r}
mod.lda <- lda(Class~.-Class01, data=train.df)
test.df$Class.lda <- predict(mod.lda, newdata=test.df)$class
with(test.df, table(Class, Class.lda))
```

```{r}
(err.lda <- with(test.df, mean(Class != Class.lda)))
```

```{r}
with(test.df, table(Class.log, Class.lda))
```

```{r}
scale.lda <- mod.lda$scaling
(scale.lda <- scale.lda/sqrt(sum(scale.lda^2)))
```

```{r}
data.mat <- data.matrix(
subset(test.df,
select=-c(Class,Class01,prob.log,Class.log,Class.lda)))
proj.vals <- data.mat %*% scale.lda
test.df$proj <- as.numeric(proj.vals)
test.df %>%
ggplot()+
geom_density(aes(proj,fill=factor(Class)))+
labs(title="Separation of Classes(opel and bus) in LDA Space (dimension=1)",
fill="Class",x="")+
facet_grid(Class~.)
```

Three Classes Repeat: LDA Only
```{r}
Vehicle3.df <- Vehicle %>%
filter(Class %in% c("opel","van", "bus")) %>%
droplevels()
```
divide the data to test and train
```{r}
N <- nrow(Vehicle3.df)
train <- sample(1:N,N/2,rep=F)
train.df <- Vehicle3.df[train,]
test.df <- Vehicle3.df[-train,]
mod.lda <- lda(Class~. , data=train.df)
test.df$Class.lda <- predict(mod.lda, newdata=test.df)$class
with(test.df, table(Class, Class.lda))
```

```{r}
(err.lda <- with(test.df, mean(Class != Class.lda)))
```

```{r}
scale.lda <- mod.lda$scaling
(scale.lda <- scale.lda/sqrt(sum(scale.lda^2)))
```

```{r}
data.mat <- data.matrix(subset(test.df, select=-c(Class,Class.lda)))
proj.vals <- data.mat %*% scale.lda
test.df$proj1 <- as.numeric(proj.vals[,1])
test.df$proj2 <- as.numeric(proj.vals[,2])
test.df %>%
ggplot()+
geom_point(aes(proj1,proj2,color=Class))+
labs(title = "Three Class Separation the Projection Space")
```

Four Classes:


Three Classes Repeat: LDA Only
divide the data to test and train
```{r}
N <- nrow(Vehicle)
train <- sample(1:N,N/2,rep=F)
train.df <- Vehicle[train,]
test.df <- Vehicle[-train,]
mod.lda <- lda(Class~., data=train.df)
test.df$Class.lda <- predict(mod.lda, newdata=test.df)$class
with(test.df, table(Class, Class.lda))
```

```{r}
(err.lda <- with(test.df, mean(Class != Class.lda)))
```

```{r}
scale.lda <- mod.lda$scaling
(scale.lda <- scale.lda/sqrt(sum(scale.lda^2)))
```

```{r}
data.mat <- data.matrix(subset(test.df, select=-c(Class,Class.lda)))
proj.vals <- data.mat %*% scale.lda
test.df$proj1 <- as.numeric(proj.vals[,1])
test.df$proj2 <- as.numeric(proj.vals[,2])
test.df$proj3 <- as.numeric(proj.vals[,3])
test.df %>%
ggplot()+
geom_point(aes(proj1,proj2,color=Class))+
labs(title = "Three Class Separation the Projection Space")
```

ggplot in 3D
```{r}
library("gg3D")
```

```{r}
cols <- c("red","blue","orange")
test.df %>%
ggplot( aes(x=proj1, y=proj2, z=proj3, color=factor(Class))) +
theme_void() +
axes_3D(theta=110,phi=10) +
scale_color_manual(values=c(cols,"cyan"))+
stat_3D()+
labs(title="3 Dimensional Projection Space",
color="Class")
```
