---
title: 'Assignment 1: Loess MSE via Bootstrap'
author: "Shan Chen"
date: "2/21/2020"
output:
  html_document: default
  pdf_document: default
  word_document: default
---

# Instructions
Recall we used a fixed data set (data.df) and then bootstrapped it to find a MSE for the loess model. In the case in class, we  used a fixed value of span (0.75 or so). 

Redo the bootstrapped MSE calculation with values of span in the interval (.1,3). Can you settle on a value of the span that minimizes the MSE? 

Plan: It might help to write a helper function that  will compute the estimated (mean) MSE and variability (variance or sd) as a function of the span. From these, create a plot that shows MSE as a function of span, along with some indication of the variability of the estimate. For example, you should plot mse, mse+2*sd, mse-2*sd all as a function of span.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,cache = TRUE)
```

```{r}
library(tidyverse)
```
Recall we used a fixed data set (data.df) and then bootstrapped it to find a MSE for the loess model. In the case in class, we  used a fixed value of span (0.75 or so). 

Replace the underlying function f(x) (=b0+b1*x) with

  f(x)= sin(x)-3*cos(3*x)

Generate the data as before...
```{r}
b0 <- 1
b1 <- 3
##f <- function(x) b0+b1*x
f <- function(x) sin(x)-3*cos(3*x)
```


```{r}
buildData <- function(N,b0,b1,sigma){
  x <- rnorm(N,0,1)
  y <-f(x)+rnorm(N,0,sigma)
  data.frame(x,y)
}

N <- 50  ##sample size
sigma <- 2.5
data.df <- buildData(N,b0,b1,sigma)
```
Display data
```{r}
ggplot(data.df)+
    geom_point(aes(x,y))+
  labs(title="Generated Data from f <- function(x) sin(x)-3*cos(3*x)")
```
## Loess model

First build and predict with a loess model to make sure we know how this works.
First with a fix span value of 0.25
```{r}
theSpan <- .25
##theSpan <- 100
mod.lo <- loess(y~x,data=data.df,span=theSpan) ##set up loess
data.df$pred <- predict(mod.lo)
mse <- with(data.df,mean((y-pred)^2)) #calculate mse
```

```{r}
data.df %>% 
ggplot()+
    geom_point(aes(x,y),color="blue")+
      geom_line(aes(x,pred),color="red")+
  labs(title="Synthetic Data from f <- function(x) sin(x)-3*cos(3*x)",
       subtitle=sprintf("With Loess  (span = %s): MSE=%s", theSpan,round(mse,3)))
```

# Bootstrapping and generate new testing data

```{r}
theSpan <- .3
N <- nrow(data.df)
bootSample <- sample(1:N,N,rep=T)
boot.df <- data.df[bootSample,]
#see test data
ggplot(boot.df)+
    geom_point(aes(x,y))+
  labs(title="Generated test Data")
#modeling and calculating loses
mod.lo <- loess(y~x,data=data.df,span=theSpan)
boot.df$pred <- predict(mod.lo,newdata=boot.df)
mse.boot <- with(boot.df,mean((y-pred)^2))
c(mse,mse.boot)
```

These two values should be close. The mse.boot has  more validity since we predicted on a different data set  from the one on which the model was built. ##But I do not see really close mse honestly.

## Repeat the Bootstrapping 100 times
```{r}
numBoots <- 100
mseVals <- array(numeric(0),numBoots)
### 
for(m in 1:numBoots){
  ## training data
 bootSample <- sample(1:N,N,rep=T)
 train.df <- data.df[bootSample,] 
 ## testing data
 bootSample <- sample(1:N,N,rep=T)
 test.df <- data.df[bootSample,] 
 ##model and predict
 mod.lo <- loess(y~x,data=train.df,span=.3)
 boot.df$pred <- predict(mod.lo,test.df)
 ##save the mse values
 mseVals[m] <- with(test.df,mean((y-pred)^2))
}

```
Now use the mean and standard deviation as a guage of the MSE estimate.

```{r}
c(mean(mseVals),sd(mseVals))
```

```{r}
mseFunc <- function(theSpan,numBoots){
  mseVals <- array(numeric(0),numBoots)
  ### 
  m <- 1
  for(m in 1:numBoots){
    ## bootstrap for training data
    bootSample <- sample(1:N,N,rep=T)
    train.df <- data.df[bootSample,] 

    ## bootstrap again for testing data
    bootSample <- sample(1:N,N,rep=T)
    test.df <- data.df[bootSample,] 
    ##model and predict
    mod.lo <- loess(y~x,data=train.df,span=theSpan)
    test.df$pred <- predict(mod.lo,newdata=test.df)
    test.df$y
    ##save the mse values
    mseVals[m] <- with(test.df,mean((y-pred)^2,na.rm=T))
  } 
  # Return the mean and sd of the mseVals
  c(mean(mseVals),sd(mseVals))
  }
```
See result
```{r}
mseFunc(5,2)
```
```{r}
mseFunc(.3,200)
```

## MSE as a function of span
Now we are ready to put all this together and plot mse (+/- 2*standard deviation).

```{r}
numSpans <- 100
numBoots <- 200
minSpan <- 0.25
maxSpan <- 2.0
spanVals <- seq(minSpan,maxSpan,length=numSpans)
## store the mean and standard deviation of the mse for each spanVal
spanMSEVals <- matrix(nrow=numSpans,ncol=2)
k <- 1
for(theSpan in spanVals){
  
  spanMSEVals[k,] <- mseFunc(theSpan,numBoots)
  print(c(theSpan,spanMSEVals[k,1]))
  k <- k+1
}
```

Display result
```{r}

data.frame(span=spanVals,mse=spanMSEVals[,1],sdev=spanMSEVals[,2]) %>% 
  ggplot()+
  geom_point(aes(span,mse),color="blue")+
    geom_point(aes(span,mse+sdev),color="red")+
      geom_point(aes(span,mse-sdev),color="red")
```

