---
title: "Linear/Quadratic Discriminant Analysis"
author: "Shan Chen"
output:
  html_document:
    df_print: paged
  pdf_document: default
---

```{r}
library(tidyverse)
library(FNN)
```

# Assignment 2: Wine Quality Prediction
Go to: https://archive.ics.uci.edu/ml/datasets/Wine+Quality
Use the white wine data set. Build a model to predict quality as
a function of the predictors. Compare linear regression with KNN
(using knn.reg)
For linear regression, use CV and/or bootstrap to determine the best (or at least a good)
set of predictors.
For KNN, determine the best choice of k.

Note: before starting the modeling, scale the predictors data before  applying knn.

```{r}
wine.df <- read_delim("winequality-white.csv",delim=";")
```
```{r}
#check data
dim(wine.df)
#simple linear model, 10folds
N <- nrow(wine.df)
numFolds <- 10
folds <- sample(1:numFolds,N,rep=T)
errs <- numeric(numFolds)
for(fold in 1:numFolds){
  train.df <- wine.df[folds != fold,]
  test.df <- wine.df[folds == fold,]
  mod.lm <- lm(quality ~ ., data=train.df)
  pred <- predict(mod.lm, newdata=test.df)
  errs[fold] <- with(test.df,mean((quality-pred)^2))
}
mse.lm <- mean(errs)

```

```{r}
totPreds <- 11
numPreds <- 2 #chose 2 pre
sampPreds <- sort(sample(1:totPreds,numPreds,rep=F))
data.df <- wine.df[,c(sampPreds,totPreds+1)]

numFolds <- 10
folds <- sample(1:numFolds,N,rep=T)
errs <- numeric(numFolds)
for(fold in 1:numFolds){
  train.df <- data.df[folds != fold,]
  test.df <- data.df[folds == fold,]
  mod.lm <- lm(quality ~ ., data=train.df)
  pred <- predict(mod.lm, newdata=test.df)
  errs[fold] <- with(test.df,mean((quality-pred)^2))
}
(mse.lm <- mean(errs))
```
func that return mse values with corresponding predictors
```{r}
RandPreds <- function(numPreds){
  sampPreds <- sort(sample(1:totPreds,numPreds,rep=F))
  data.df <- wine.df[,c(sampPreds,totPreds+1)]
  folds <- sample(1:numFolds,N,rep=T)
  errs <- numeric(numFolds)
  for(fold in 1:numFolds){
    train.df <- data.df[folds != fold,]
    test.df <- data.df[folds == fold,]
    mod.lm <- lm(quality ~ ., data=train.df)
    pred <- predict(mod.lm, newdata=test.df)
    errs[fold] <- with(test.df,mean((quality-pred)^2))
  }
  mse.lm <- mean(errs)
  c(mse.lm,sampPreds)
}
```
try it out and compare results:
```{r}
RandPreds(6)
RandPreds(6)
RandPreds(6)
RandPreds(7)
RandPreds(7)
RandPreds(7)
RandPreds(8)
RandPreds(8)
RandPreds(8)

```

```{r}
RandPreds(9)
RandPreds(9)
RandPreds(10)
RandPreds(10)
RandPreds(11)
msewin.lm = RandPreds(11)
```
As we can see from result above, it can get close, but beating the full set predictors is pretty hard.


#KNN:
```{r}
wine.mat <- data.matrix(wine.df)
##just the predictors, scaled
wine.x <- scale(wine.mat[,-12])
## the response
wine.y <- wine.mat[,12]
## check that these are close to 0
colMeans(wine.x)

```

```{r}
kVal <- 10
## build the folds
numFolds <- 10
folds <- sample(1:numFolds,N,rep=T)
## Ready to cross-validate
errs <- numeric(numFolds)
for(fold in 1:numFolds){
  train.x <- wine.x[folds != fold,]
  test.x <- wine.x[folds == fold,]
  train.y <- wine.y[folds != fold]
  test.y <- wine.y[folds == fold]
  mod.knn <- knn.reg(train.x,test.x,train.y,k=kVal)
  pred <- mod.knn$pred
  errs[fold] <- mean((test.y-pred)^2)
}
mse.knn <- mean(errs)
##compare with lm....
c(mse.lm,mse.knn)
```

#repeat KNN
```{r}
mseKNN <- function(kVal){
  folds <- sample(1:numFolds,N,rep=T)
  errs <- numeric(numFolds)
  for(fold in 1:numFolds){
    train.x <- wine.x[folds != fold,]
    test.x <- wine.x[folds == fold,]
    train.y <- wine.y[folds != fold]
    test.y <- wine.y[folds == fold]
    mod.knn <- knn.reg(train.x,test.x,train.y,k=kVal)
    pred <- mod.knn$pred
    errs[fold] <- mean((test.y-pred)^2)
  }
  mean(errs)
}
```

#set k= 1-61 only odd 
```{r}
maxK <- 30
#only odd values
kVals <- 2*(1:maxK)+1
## this takes a moment or two...
errsKNN <- map_dbl(kVals,mseKNN)
```
#knn vs err by values
```{r}
data.frame(k=kVals,err=errsKNN) %>%
ggplot()+
geom_point(aes(k,err))+
geom_line(aes(k,err))
```

#findmin Kval:
#KNN won by 6%
```{r}
minid <- which.min(errsKNN)
k0id <- kVals[minid]
c(k0id)
(mse.knn <- errsKNN[k0id])
c(msewin.lm,mse.knn)
```



